### 2023-06-11 presentation/podc-presentation.tex
Added related work and final slide

\documentclass[compress]{beamer}

\usepackage{color}
\usepackage{amssymb}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage[normalem]{ulem} % for striking through text

\mode<presentation>
{
  \usetheme{Warsaw}
  \usecolortheme{beaver}
  \usefonttheme{structurebold}

  % or ...

  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}
\logo{\includegraphics[width=15mm]{YorkLogo.pdf}}

\newcommand{\Red}[1]{{\color{red}#1}}
\newcommand{\Blue}[1]{{\color{blue}#1}}
\newcommand{\Green}[1]{{\color{green}#1}}

\newcommand{\op}[1]{\mbox{\sc #1}}
\newcommand{\tb}{\hspace*{10mm}}

\renewcommand{\log}{\mbox{log }} % font looks funny for usual \log

\newcommand{\ignore}[1]{}

\title{A Wait-Free Queue with Polylogarithmic Step Complexity}

\author{Hossein Naderibeni \and Eric Ruppert}
\date{June 21, 2023}

\begin{document}

\begin{frame}
\maketitle
\end{frame}

\begin{frame}{Lock-Free Queue using CAS}
Michael and Scott Queue [PODC 1996]

\bigskip

% We store the elements of the queue in a singly-linked list.
% Enqueues add elements at the left end and dequeues remove elements from the right end.
% Each enqueue creates new node and performs a CAS on the tail node's next pointer to add it to the left end (and then advances the Tail pointer)
% Each dequeue uses a CAS to advance the Head pointer

\only<1>{\input{MSQueue-48to50.pdf_t}}
\only<2>{\input{MSQueue-48to51.pdf_t}}
\only<3>{\input{MSQueue-48to52.pdf_t}}
\only<4>{\input{MSQueue-48-50to53.pdf_t}}
\only<5>{\input{MSQueue-50to54.pdf_t}}
\end{frame}

\begin{frame}{CAS Retry Problem}
Suppose $p$ processes want to enqueue simultaneously.
\medskip

\only<1>{\input{CASretry-49to50.pdf_t}}
\only<2>{\input{CASretry-49to51.pdf_t}}
\only<3->{\input{CASretry-49to52.pdf_t}}

\medskip
\visible<4->{
\begin{itemize}
\item one CAS of next pointer succeeds.
\item other \op{Enqueues} must try again.
\item<5> $\Rightarrow$ starvation and $\Omega(p)$ steps per operation (amortized)
\end{itemize}
}
\end{frame}

\begin{frame}{Other Lock-Free Queues}

Other \Red{list-based} queues
\begin{itemize}
\item
add elimination array \hfill {\small [Moir et al.\ 2005]}
\item
baskets queue \hfill {\small [Hoffman, Shalev, Shavit 2007]}
\item
doubly-linked list + optimism \hfill {\small [Ladan-Mozes, Shavit 2008]}
\item
fast path, slow path \hfill {\small [Kogan, Petrank 2012]}
\item
futures \hfill {\small [Kogan, Herlihy 2014]}
\end{itemize}
\visible<2->{All have CAS retry problem  \hfill {\small [Morrisson, Afek 2013]}}

\smallskip

\visible<3->{So do \Red{array-based} queues

\smallskip

All$^*$ previous queues take amortized $\Omega(p)$ steps per operation}

\end{frame}

\begin{frame}{$^*$Exceptions: Sublinear Time Queues}

\begin{block}{Restricted queues}
\begin{itemize}
\item
1 enqueuer, multiple dequeuers \hfill {\small [David 2004]}
\item
1 dequeuer, multiple enqueuers \hfill {\small [Jayanti, Petrovic 2005]}
\end{itemize}
\end{block}

\pause

\begin{block}{Other primitives}
\begin{itemize}
\item
$O(\sqrt{p})$ using unusual double-word RMW instructions
 \hfill {\small [Khanchandani, Wattenhofer 2018]}
\end{itemize}
\end{block}

\pause

\begin{block}{Universal constructions}
\begin{itemize}
\item
$O(\log p)$ using huge words\\
\mbox{ }\hfill{\small [Afek, Dauber, Touitou 1995; Jayanti 1998]}
\item
$\Omega(p)$ with reasonably-sized words
\end{itemize}
\end{block}
\end{frame}

\begin{frame}{Lower Bound}

All previous multi-enqueuer, multi-dequeuer queues take\\
$\Omega(p)$ steps per operation.

\medskip

For many  data structures, fastest lock-free operations take
$O(sequential\ complexity + contention)$ steps

\pause

\begin{block}{Lower Bound \hfill {\small [Attiya, Fouren 2017]}}
\begin{itemize}
\item
Amortized step complexity for any bag is $\Omega(contention)$.
\pause
\item
But lower bound holds only if $contention$ is $O(\log\log p)$
\end{itemize}
\end{block}

\end{frame}

\begin{frame}{Our Result}

\begin{block}{New Queue}
\begin{itemize}
\item $O(\log p)$ steps per \op{Enqueue}
\item $O(\log^2 p + \log q)$ steps per \op{Dequeue}
\item wait-free
\item uses CAS on reasonable-size words
\item bounded space version: $O(\log p\ \log(p+q))$ amortized steps per operation
(relies on safe GC)
\end{itemize}
\end{block}

$p = \#$ processes\\
$q = \#$ elements in queue

\medskip

Previous lock-free queues take $\Omega(p)$ steps per operation.

\end{frame}

\begin{frame}{Future Directions}

\begin{itemize}
\item
Practical implementation\\
(perhaps slow path of fast path slow path method)
\item
Extend technique to other data structures\\
 (stacks and deques are recently done)
 \item
Close gap between\\
$\Omega(\log \log p)$ lower bound \hfill {\small [Attiya Fouren 2017]}\\
$O(\log^2 p + \log q)$ upper bound \hfill {\small [this work]}

\end{itemize}

\end{frame}

\end{document}